Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-06-09 manuscript/manuscript.tex
updates ms fig 2 to show differences from mean cover, matches better with sensitivity analysis

\begin{align}
x_{jq,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{G}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) T[0,1]
\end{align}

where $x_{jq,t}$ is the log of species' $j$ proportional cover in
quadrat $q$ at time $t$ and all other parameters are as in the
individual-level growth model (Eq. \#). The log normal likelihood
includes a truncation ($T${[}0,1{]}) to ensure that predicted values do
not exceed 100\% cover.

@@

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. We recignize
such short MCMC chains may surprise those more familiar with other MCMC
samplers (i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally
efficient, which reduces the number of iterations needed to achieve
convergence. We assessed convergence visually and made sure scale
reduction factors for all parameters were less than 1.01. For the
purposes of introducing stochasticity in our population models, we saved
the final 1,000 iterations from each of the three MCMC chains for all
parameters to be used as randomly drawn values during population
simulation. Importantly, this step alleviates the need to reduce model
parameters since sampling from the full parameter space in the MCMC
ensures that if a parameter broadly overlaps zero, on average the effect
in the population models will also be near zero.

@@

We performed one-step-ahead forecasts to compare the two models (IPM
vs.~QBM). We did not use the random year effects for these forecasts.
Thus, beyond the effect of genet size or quadrat cover, only the climate
covariates could affect expect plant cover. The models had similar
accuracy ($\rho$, correlation between observations and mean prediction
from 100 forecasts) across all species (Table 1). However, the IPM had
significantly lower overall error (MAE, mean absolute error) for two
species (\emph{H. comita} and \emph{P. smithii}), and in no case did the
QBM significantly outperform the IPM (Table 1).

@@

The IPM and QBM produced consistent predictions for \emph{H. comita}
under increased precipitation and when both precipitation and
temperature were increased (Figure 3). However, the IPM predicted more
modest changes than the QBM, and for a temperature increase the two
models differed: the IPM predicted an increase in cover while the QBM
predicted the opposite.

@@

>>>>

\begin{align}
x_{jq,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{G}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) \text{T}[0,1]
\end{align}

where $x_{jq,t}$ is the log of species' $j$ proportional cover in
quadrat $q$ at time $t$ and all other parameters are as in the
individual-level growth model (Eq. \#). The log normal likelihood
includes a truncation (T{[}0,1{]}) to ensure that predicted values do
not exceed 100\% cover.

@@

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. We recignize
such short MCMC chains may surprise those more familiar with other MCMC
samplers (i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally
efficient, which reduces the number of iterations needed to achieve
convergence. We assessed convergence visually and made sure scale
reduction factors for all parameters were less than 1.01. For the
purposes of including parameter uncertaintyy in our population models,
we saved the final 1,000 iterations from each of the three MCMC chains
for all parameters to be used as randomly drawn values during population
simulation. This step alleviates the need to reduce model parameters by
model selection since sampling from the full parameter space in the MCMC
ensures that if a parameter broadly overlaps zero, on average the effect
in the population models will also be near zero.

@@

We performed one-step-ahead forecasts to compare the two models (IPM
vs.~QBM). We did not use the random year effects for these forecasts.
Thus, beyond the effect of genet size or quadrat cover, only the climate
covariates could affect expect plant cover. The IPM had significantly
lower overall error (MAE, mean absolute error) for two species (\emph{H.
comata} and \emph{P. smithii}; Table 1). In no case did the QBM
significantly outperform the IPM (Table 1). As a metric of model
uncertainty, we calculated the euclidean distance between the upper and
lower 90\% quantiles for each model's forecasts. The IPM always
outperformed the QBM in this respect (Table 1), with consistently lower
distances between the 90\% quantiles. In general the IPM outperformed
the QBM because it had (i) lower MAE for two species, (ii) statistically
similar MAE for the other two species, and (iii) considerably more
precise forecasts for all species.

@@

The IPM and QBM produced consistent predictions for \emph{H. comata}
under increased precipitation and when both precipitation and
temperature were increased (Figure 3). However, the IPM predicted more
modest changes than the QBM, and for a temperature increase the two
models differed: the IPM predicted an increase in cover while the QBM
predicted the opposite.

@@

.
